---
title: "Dynamic Model"
author: "Alex Hoagland"
date: "`r Sys.Date()`"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
# Note: this model matches on the intensive margin only 
knitr::opts_chunk$set(echo = FALSE)

set.seed(20250212) # set seed for random draws

library(tidyverse) # call the relevant library
library(gmm) # For GMM estimation
library(EnvStats) # for type 1 distribution
library(broom) # for RD regression
library(purrr) # for parallelizing
library(furrr) # for parallelizing
library(pracma) # for numerical integration
library(Rcpp) # for CS loops (to go faster?)
library(future)
library(parallel)

# path <- "C:\\Users\\pc\\Dropbox\\Spouses in Medicare\\Model\\" # for Columbus machine
path <- "C:\\Users\\alexh\\Dropbox\\Spouses in Medicare\\Model\\" # TO machine
```

## Model Estimation (left out of slide compiling)
-     Note that Eliason et al. used the first set of moments to construct their likelihood function
-     By using GMM, we allow for the additional moment of the RD estimate. 
-     The parameters to estimate are the $\lambda$ parameters, $\alpha$, $\omega$, and $\sigma_\varepsilon$.

```{r gmm-estimation,  eval=FALSE}
# Parameters to specify
N <- 5000 # how many individuals to sample
treated <- TRUE # are we doing this for treated group or control group? 
# note that per Eliason et al., delta should be 0.99986 for daily rate when annual discount rate is 0.95. So it's basically 1

# A needed function to use a given person's shocks to return outcomes
simulate_person <- function(p, simdata, timevars, omega, myscale) { 
  # Local copy with that person's shocks and inflate by the scale
  local_timevars <- simdata %>% filter(id == p) %>% select(eps_d, eps_k, t) %>% 
    left_join(timevars, by = "t") %>% 
    mutate(eps_d = myscale*eps_d, 
           eps_k = myscale*eps_k, # transform shocks accordingly
           EV = ifelse(t == 100, omega, NA))
  
  for (d in 99:1) { # backward loop
    local_timevars$EV[local_timevars$t == d] <- 
      local_timevars$u[local_timevars$t == d + 1] + 
      log(exp(1 * local_timevars$EV[local_timevars$t == d + 1]) + 1)
  }
  local_timevars$V <- NA
  local_timevars$V[local_timevars$t == 100] <- local_timevars$u[local_timevars$t == 100]
  
  for (d in 99:1) { # backward loop
    local_timevars$V[local_timevars$t == d] <- 
      local_timevars$u[local_timevars$t == d] + 
      pmax(local_timevars$eps_k[local_timevars$t == d] + 1 * local_timevars$EV[local_timevars$t == d + 1],
           local_timevars$eps_d[local_timevars$t == d])
  }
  
  # Calculate cumulative value
  local_timevars <- local_timevars %>% mutate(cumV = cumsum(V))
  
  # Find the day that maximizes cumulative value
  exit_day <- which.max(local_timevars$cumV)
  
  # If value on the first day is negative, then return 0 instead
  # exit_day <- ifelse(local_timevars$V[1] < 0, 0, exit_day)
  
  # If value is infinite, return corner solution
  exit_day <- ifelse(local_timevars$V[1] == Inf, 100, exit_day)
  return(exit_day)
}

# Load observed moments
observed_moments <- read_csv(paste0(path, "ObservedMoments.csv", sep="")
names(observed_moments)[1] <- "exit_day"

# Change moments to intensive margins
observed_moments <- observed_moments %>% 
  filter(exit_day > 0) %>% 
  mutate(frac_observed_treated = frac_observed_treated / sum(frac_observed_treated),
         frac_observed_control = frac_observed_control / sum(frac_observed_control))

 # moments need the average, median, and SD of LOS for both treated and control groups 
  if (treated == TRUE) {
    mean_cond_treat <- sum(observed_moments$exit_day*observed_moments$frac_observed_treated)
    med_cond_treat <- observed_moments$exit_day[which.max(cumsum(observed_moments$frac_observed_treated) >= 0.5)]
    sd_cond_treat <- sqrt(sum(observed_moments$frac_observed_treated * (observed_moments$exit_day - mean_cond_treat)^2))
   
    rd_moment <- observed_moments %>%
      mutate(past_cutoff = ifelse(exit_day > 21, 1, 0)) %>%
      filter(abs(exit_day - 22) < 11.22, exit_day != 22) %>%
      mutate(
        distance = exit_day - 22,
        weight = 1 - abs(distance) / 11.22
      ) %>%
      group_map(~ {
        model <- lm(frac_observed_treated ~ past_cutoff, data = .x, weights = weight)
        coef(model)["past_cutoff"]
      }) %>%
      unlist()  # if you want a plain numeric vector
    
  } else {
    mean_cond_control <- sum(observed_moments$exit_day*observed_moments$frac_observed_control)
    med_cond_control <- observed_moments$exit_day[which.max(cumsum(observed_moments$frac_observed_control) >= 0.5)]
    sd_cond_control <- sqrt(sum(observed_moments$frac_observed_control * (observed_moments$exit_day - mean_cond_control)^2))
    
    rd_moment <- observed_moments %>%
      mutate(past_cutoff = ifelse(exit_day > 21, 1, 0)) %>%
      filter(abs(exit_day - 22) < 11.22, exit_day != 22) %>%
      mutate(
        distance = exit_day - 22,
        weight = 1 - abs(distance) / 11.22
      ) %>%
      group_map(~ {
        model <- lm(frac_observed_control ~ past_cutoff, data = .x, weights = weight)
        coef(model)["past_cutoff"]
      }) %>%
      unlist()  # if you want a plain numeric vector

  }

# make a dataset of N individuals with unique shocks 
simdata <- expand_grid(
  id = 1:N,
  t = 1:100
) %>%
  arrange(id, t) %>% 
  mutate(eps_k = revd(n = (100*N), scale=1), 
         eps_d = revd(n = (100*N), scale=1)) 
  # note: we aren't using any individual characteristics here, so these can be simulated individuals who differ only in their shocks 

# Moment function 
simulate_moments <- function(structparams, data=simdata) {
  cat("Testing parameters:", structparams, "\n")  # to see progress
  
  l0 <- structparams[1] #scaling
  l1 <- structparams[2]/1000 #scaling
  l2 <- structparams[3]/1000 #scaling
  l3 <- structparams[4]/1000 # scaling
  alpha <- structparams[5]
  omega <- structparams[6]
  
  # make a dataframe with 100 periods and a variable t
  timevars <- data.frame(t = 1:100) %>%
    mutate(lambda = l0 + l1*t + l2*t^2 + l3*t^3) %>%
    # mutate(ubump = ifelse(t == 21, structparams[8], 
    #                       ifelse(t == 22, structparams[9], 
    #                              ifelse(t == 23, structparams[10], 0)))) %>%
    mutate(p = ifelse(t < 22, 1e-5, 37.5),
           omega = omega,
           delta = 1,
           alpha = alpha,
           u = lambda + alpha*p ) # + ubump)
  
  # solve the model for each individual with different shocks in parallel
   pred_moments <- unique(simdata$id) %>% future_map_int(
    ~simulate_person(.x, simdata, timevars, omega, myscale=structparams[7]),
    .options = furrr_options(seed = TRUE)
  )
  pred_moments <- tibble(id = unique(simdata$id), exit_day = pred_moments)
   
  # Then you can calculate the moments
  moments <- pred_moments %>% group_by(exit_day) %>% summarize(n = n()) %>% 
    mutate(p = n/N) %>% # fill in with 0s if needed
    right_join(data.frame(exit_day = 1:100), by = "exit_day") %>% 
    mutate(p = ifelse(is.na(p), 0, p))
  
  rd_moment1 <- moments %>%
      mutate(past_cutoff = ifelse(exit_day > 21, 1, 0)) %>%
      filter(abs(exit_day - 22) < 11.22, exit_day != 22) %>%
      mutate(
        distance = exit_day - 22,
        weight = 1 - abs(distance) / 11.22
      ) %>%
      group_map(~ {
        model <- lm(p ~ past_cutoff, data = .x, weights = weight)
        coef(model)["past_cutoff"]
      }) %>%
      unlist()  # if you want a plain numeric vector
  
  # Now construct moments as the difference between these outputs and the actual moments in Medicare data
  if (treated == TRUE) { 
    moments <- moments %>% left_join(observed_moments, by = "exit_day") %>% 
    mutate(moment = abs(p - frac_observed_treated)) %>% filter(exit_day <= 40) %>% 
      complete(exit_day = 1:40, fill = list(moment = 0))  # ensures exactly 40 elements
      
    # ggplot(moments, aes(x=exit_day)) +
    #   geom_point(aes(y=p, color="Predicted"), size=3) +
    #   geom_point(aes(y=frac_observed_control, color="Actual"), size=3) +
    #   theme_minimal() +
    #   labs(x="Day in SNF", y = "%", subtitle="Sick Spouse Group", color=NULL) +
    #   geom_vline(xintercept = 21.5, linetype="dashed") +
    #   theme(legend.position="inside",
    #       legend.position.inside = c(0.01, 0.99),       # top-left corner inside the plot area
    #       legend.justification = c("left", "top"))
    
    # ggplot(moments, aes(x=exit_day)) + geom_point(aes(y=p),color='skyblue') +
    #   geom_point(aes(y=frac_observed_treated),color="maroon")
    # ggplot(moments, aes(x=exit_day)) + geom_point(aes(y=moment),color='forestgreen') + 
    #   geom_hline(yintercept = 0)
  }
  else {
  moments <- moments %>% left_join(observed_moments, by = "exit_day") %>% 
    mutate(moment = abs(p - frac_observed_control)) %>% filter(exit_day <= 40) %>%
    complete(exit_day = 1:40, fill = list(moment = 0))  # ensures exactly 40 elements

  # ggplot(moments, aes(x=exit_day)) +
  #   geom_point(aes(y=p, color="Predicted"), size=3) +
  #   geom_point(aes(y=frac_observed_control, color="Actual"), size=3) +
  #   theme_minimal() +
  #   labs(x="Day in SNF", y = "%", subtitle="Healthy Spouse Group", color=NULL) +
  #   geom_vline(xintercept = 21.5, linetype="dashed") +
  #   theme(legend.position="inside",
  #       legend.position.inside = c(0.01, 0.99),       # top-left corner inside the plot area
  #       legend.justification = c("left", "top"))
  #  ggplot(moments, aes(x=exit_day)) +
  #    geom_point(aes(y=moment), color='forestgreen') +
  #    geom_hline(yintercept = 0)
  }
  
  moments <- moments %>% pull(moment)
  moments[41] <- abs(rd_moment-rd_moment1)
  return(moments)
}

moment_function <- function(structparams, data=simdata) {
  # Wrap the call to compute_f in tryCatch
  result <- tryCatch({
    # Attempt to compute the moments normally
    mythetavec <- structparams 
    simulate_moments(mythetavec, data=simdata)
  }, error = function(e) {
    # In case of an error, return a penalty value.
    penalty_value <- 1e10
    rep(penalty_value, 41) # change to reflect number of moments we have
  })
  
  return(result)
}
###########################################################


##### Estimate using GMM #####
# structparams[1] -- structparams[4] are the parameters for lambda_t 
# structparams[5] is alpha (should be negative) 
# structparams[6] is omega (terminal value of SNF at day 100) 
# structparams[7] governs the variance of the shocks 

# Create a weighting matrix that upweights moments in the RD bandwidth and the RD estimate
myw <- diag(41)
myw[41,41] <- 5
for(i in 1:40) { 
  myw[i,i] <- 1-abs(i-22)/22 # triangular weights in distribution
  }

# Set number of cores to parallelize
plan(multisession, workers = parallel::detectCores() - 1)

scale_ratio <- c(5, 30, 5, .05, .001, 1, 10) # not sure what this should look like at first
myouts <- gmm(moment_function, 
              x = simdata,
              # weightsMatrix=as.matrix(myw), # weight moments close to bandwidth higher
              # method = "L-BFGS-B",       # Spscify optimizer method supporting bounds if needed
              t0 = c(-5.7, 28.9, -5.39, 0.0589, -.01, -1.72, 8.09), # starting values,
              crit = 1e-12, # tolerance level (default = 1e-7) 
              onlyCoefficients = TRUE, # don't need standard errors for the moment
              control = list(parscale = scale_ratio, trace = 1)) # to handle differently scaled parameters 

# save file once computed 
if (treated == TRUE) { 
  save.image(paste0(path, "Convergence_DynamicModel_TREATED_20250410.RData", sep=""))
} else { 
  save.image(paste0(path, "Convergence_DynamicModel_CONTROL_20250410.RData", sep=""))
}
###########################################################
```


## Model Fit: Sick Spouse Group
```{r model-fit-treated}
# Treated figure
rm(list = setdiff(ls(), "path"))
load(paste0(path, "Convergence_DynamicModel_TREATED_20250410.RData", sep=""))
structparams <- myouts$coefficients

simdata <- expand_grid(
  id = 1:N,
  t = 1:100
) %>%
  arrange(id, t) %>% 
  mutate(eps_k = revd(n = (100*N), scale=1), 
         eps_d = revd(n = (100*N), scale=1)) 

  l0 <- structparams[1] #scaling
  l1 <- structparams[2]/1000 #scaling
  l2 <- structparams[3]/1000 #scaling
  l3 <- structparams[4]/1000 # scaling
  alpha <- structparams[5]
  omega <- structparams[6]
  
  # make a dataframe with 100 periods and a variable t
  timevars <- data.frame(t = 1:100) %>%
    mutate(lambda = l0 + l1*t + l2*t^2 + l3*t^3) %>%
    # mutate(ubump = ifelse(t == 21, structparams[8], 
    #                       ifelse(t == 22, structparams[9], 
    #                              ifelse(t == 23, structparams[10], 0)))) %>%
    mutate(p = ifelse(t < 22, 1e-5, 37.5),
           omega = omega,
           delta = 1,
           alpha = alpha,
           u = lambda + alpha*p) # + ubump)
  
  # solve the model for each individual with different shocks in parallel
   pred_moments <- unique(simdata$id) %>% future_map_int(
    ~simulate_person(.x, simdata, timevars, omega, myscale=structparams[7]),
    .options = furrr_options(seed = TRUE)
  )
  pred_moments <- tibble(id = unique(simdata$id), exit_day = pred_moments)
   
  # Then you can calculate the moments
  moments <- pred_moments %>% group_by(exit_day) %>% summarize(n = n()) %>% 
    mutate(p = n/N) %>% # fill in with 0s if needed
    right_join(data.frame(exit_day = 1:100), by = "exit_day") %>% 
    mutate(p = ifelse(is.na(p), 0, p))
  
  moments <- moments %>% left_join(observed_moments, by = "exit_day") %>% 
    mutate(moment = abs(p - frac_observed_treated)) %>% filter(exit_day <= 50) %>%
    mutate(frac_observed_treated = frac_observed_treated*100, 
           p = p * 100)
    
modelfit <- ggplot(moments, aes(x=exit_day)) +
      geom_point(aes(y=p, color="Predicted"), size=3) +
      geom_point(aes(y=frac_observed_treated, color="Actual"), size=3) +
      theme_minimal() +
      labs(x="Day in SNF", y = "%", color=NULL) +
      scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
      geom_vline(xintercept = 21.5, linetype="dashed") +
         # tuck the legend into the topŌĆÉleft of the panel
      theme(
        legend.position = c(0.02, 0.98),            # x, y in [0,1] (0 = left/bottom, 1 = right/top)
        legend.justification = c("left", "top"),          # anchor the legendŌĆÖs (0,1) corner
      legend.background = element_rect(
                           fill   = "white",        # so it doesnŌĆÖt obscure points
                           colour = "grey50",       # a little border helps
                           size   = 0.5
                         ))
      # subtitle="Sick Spouse Group",
modelfit
ggsave(modelfit, filename = paste0(path, "ModelFit_TreatmentGroup_20250530.png", sep=""), 
       width = 8, height = 6, dpi = 300)
ggsave(modelfit, filename = paste0(path, "ModelFit_TreatmentGroup_20250530.pdf", sep=""), 
       width = 8, height = 6, dpi = 300)
```

## Model Fit: Healthy Spouse Group
```{r model-fit-control}
# Treated figure
rm(list = setdiff(ls(), "path"))
load(paste0(path, "Convergence_DynamicModel_CONTROL_20250410.RData", sep=""))
structparams <- myouts$coefficients

simdata <- expand_grid(
  id = 1:N,
  t = 1:100
) %>%
  arrange(id, t) %>% 
  mutate(eps_k = revd(n = (100*N), scale=1), 
         eps_d = revd(n = (100*N), scale=1)) 

  l0 <- structparams[1] #scaling
  l1 <- structparams[2]/1000 #scaling
  l2 <- structparams[3]/1000 #scaling
  l3 <- structparams[4]/1000 # scaling
  alpha <- structparams[5]
  omega <- structparams[6]
  
  # make a dataframe with 100 periods and a variable t
  timevars <- data.frame(t = 1:100) %>%
    mutate(lambda = l0 + l1*t + l2*t^2 + l3*t^3) %>%
    # mutate(ubump = ifelse(t == 21, structparams[8], 
    #                       ifelse(t == 22, structparams[9], 
    #                              ifelse(t == 23, structparams[10], 0)))) %>%
    mutate(p = ifelse(t < 22, 1e-5, 37.5),
           omega = omega,
           delta = 1,
           alpha = alpha,
           u = lambda + alpha*p) # + ubump)
  
  # solve the model for each individual with different shocks in parallel
   pred_moments <- unique(simdata$id) %>% future_map_int(
    ~simulate_person(.x, simdata, timevars, omega, myscale=structparams[7]),
    .options = furrr_options(seed = TRUE)
  )
  pred_moments <- tibble(id = unique(simdata$id), exit_day = pred_moments)
   
  # Then you can calculate the moments
  moments <- pred_moments %>% group_by(exit_day) %>% summarize(n = n()) %>% 
    mutate(p = n/N) %>% # fill in with 0s if needed
    right_join(data.frame(exit_day = 1:100), by = "exit_day") %>% 
    mutate(p = ifelse(is.na(p), 0, p))
  
  moments <- moments %>% left_join(observed_moments, by = "exit_day") %>% 
    mutate(moment = abs(p - frac_observed_control)) %>% filter(exit_day <= 50) %>% 
    mutate(p = p*100, frac_observed_control = frac_observed_control*100)
      
modelfit <- ggplot(moments, aes(x=exit_day)) +
      geom_point(aes(y=p, color="Predicted"), size=3) +
      geom_point(aes(y=frac_observed_control, color="Actual"), size=3) +
      theme_minimal() +
      labs(x="Day in SNF", y = "%", color=NULL) +
      scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
      geom_vline(xintercept = 21.5, linetype="dashed") +
         # tuck the legend into the topŌĆÉleft of the panel
      theme(
        legend.position = c(0.02, 0.98),            # x, y in [0,1] (0 = left/bottom, 1 = right/top)
        legend.justification = c("left", "top"),          # anchor the legendŌĆÖs (0,1) corner
      legend.background = element_rect(
                           fill   = "white",        # so it doesnŌĆÖt obscure points
                           colour = "grey50",       # a little border helps
                           size   = 0.5
                         ))
      # subtitle="Healthy Spouse Group",
modelfit
ggsave(modelfit, filename = paste0(path, "ModelFit_ControlGroup_20250530.png", sep=""), 
       width = 8, height = 6, dpi = 300)
ggsave(modelfit, filename = paste0(path, "ModelFit_ControlGroup_20250530.pdf", sep=""), 
       width = 8, height = 6, dpi = 300)
```

## Extensive Margin Shifts
-    Calculate average u(day 1) in both groups
-    Add shift parameter so that the fraction with positive utility matches extensive margin data.

```{r ext-margin}
# Load estimated parameters from both estimations 
rm(list = setdiff(ls(), "path"))
load(paste0(path, "Convergence_DynamicModel_TREATED_20250410.RData", sep=""))
sp_treated <- myouts$coefficients
rm(list = setdiff(ls(), c("path", "sp_treated")))
load(paste0(path, "Convergence_DynamicModel_CONTROL_20250410.RData", sep=""))
sp_control <- myouts$coefficients
tokeep <- c("path", "sp_treated", "sp_control", "moment_function", "simulate_person", "simulate_moments")
rm(list = setdiff(ls(), tokeep))

# Load observed moments
observed_moments <- read_csv(paste0(path, "ObservedMoments.csv", sep=""))
names(observed_moments)[1] <- "exit_day"
observed_moments <- observed_moments %>% 
  mutate(group = exit_day >= 1) %>% 
  group_by(group) %>% 
  summarize(frac_observed_control = sum(frac_observed_control), 
            frac_observed_treated = sum(frac_observed_treated))
  # extensive margin differences

# Now calculate the average utility on day 1 in both groups
# first, control group
N <- 5000
simdata <- expand_grid(
  id = 1:N,
  t = 1:100
) %>%
  arrange(id, t) %>% 
  mutate(eps_k = revd(n = (100*N), scale=1), 
         eps_d = revd(n = (100*N), scale=1)) 

l0 <- sp_control[1] #scaling
l1 <- sp_control[2]/1000 #scaling
l2 <- sp_control[3]/1000 #scaling
l3 <- sp_control[4]/1000 # scaling
alpha <- sp_control[5]
omega <- sp_control[6]

# make dataframe 
timevars <- data.frame(t = 1:100) %>%
  mutate(lambda = l0 + l1*t + l2*t^2 + l3*t^3) %>%
  mutate(p = ifelse(t < 22, 1e-5, 37.5),
         omega = omega,
         delta = 1,
         alpha = alpha,
         u = lambda + alpha*p) # + ubump)

# A needed function to use a given person's shocks to return outcomes
valueday1 <- function(p, simdata, timevars, omega, myscale) { 
  # Local copy with that person's shocks and inflate by the scale
  local_timevars <- simdata %>% filter(id == p) %>% select(eps_d, eps_k, t) %>% 
    left_join(timevars, by = "t") %>% 
    mutate(eps_d = myscale*eps_d, 
           eps_k = myscale*eps_k, # transform shocks accordingly
           EV = ifelse(t == 100, omega, NA))
  
  for (d in 99:1) { # backward loop
    local_timevars$EV[local_timevars$t == d] <- 
      local_timevars$u[local_timevars$t == d + 1] + 
      log(exp(1 * local_timevars$EV[local_timevars$t == d + 1]) + 1)
  }
  local_timevars$V <- NA
  local_timevars$V[local_timevars$t == 100] <- local_timevars$u[local_timevars$t == 100]
  
  for (d in 99:1) { # backward loop
    local_timevars$V[local_timevars$t == d] <- 
      local_timevars$u[local_timevars$t == d] + 
      pmax(local_timevars$eps_k[local_timevars$t == d] + 1 * local_timevars$EV[local_timevars$t == d + 1],
           local_timevars$eps_d[local_timevars$t == d])
  }
  
  # Calculate cumulative value
  local_timevars <- local_timevars %>% mutate(cumV = cumsum(V))
  day1 <- local_timevars %>% filter(t == 1) %>% select(cumV) %>% as.numeric()
  return(day1)
}

# solve the model for each individual with different shocks in parallel
 pred_moments <- unique(simdata$id) %>% future_map_dbl(
  ~valueday1(.x, simdata, timevars, omega, myscale=sp_control[7]),
  .options = furrr_options(seed = TRUE)
)
pred_moments <- tibble(id = unique(simdata$id), value_day1 = pred_moments)

# now we want to shift these up and down so that extensive margin moment matches 
# TODO: do we want to include this as a parameter in the GMM estimation above? Can include for the next run
shift <- quantile(pred_moments$value_day1,1-0.2141) # quantile(pred_moments$value_day1,observed_moments$frac_observed_control[1])
pred_moments$value_day1 <- pred_moments$value_day1 - shift
sp_control <- c(sp_control, shift)

# now, treated group
l0 <- sp_treated[1] #scaling
l1 <- sp_treated[2]/1000 #scaling
l2 <- sp_treated[3]/1000 #scaling
l3 <- sp_treated[4]/1000 # scaling
alpha <- sp_treated[5]
omega <- sp_treated[6]

# make dataframe 
timevars <- data.frame(t = 1:100) %>%
  mutate(lambda = l0 + l1*t + l2*t^2 + l3*t^3) %>%
  mutate(p = ifelse(t < 22, 1e-5, 37.5),
         omega = omega,
         delta = 1,
         alpha = alpha,
         u = lambda + alpha*p) # + ubump)

# A needed function to use a given person's shocks to return outcomes
valueday1 <- function(p, simdata, timevars, omega, myscale) { 
  # Local copy with that person's shocks and inflate by the scale
  local_timevars <- simdata %>% filter(id == p) %>% select(eps_d, eps_k, t) %>% 
    left_join(timevars, by = "t") %>% 
    mutate(eps_d = myscale*eps_d, 
           eps_k = myscale*eps_k, # transform shocks accordingly
           EV = ifelse(t == 100, omega, NA))
  
  for (d in 99:1) { # backward loop
    local_timevars$EV[local_timevars$t == d] <- 
      local_timevars$u[local_timevars$t == d + 1] + 
      log(exp(1 * local_timevars$EV[local_timevars$t == d + 1]) + 1)
  }
  local_timevars$V <- NA
  local_timevars$V[local_timevars$t == 100] <- local_timevars$u[local_timevars$t == 100]
  
  for (d in 99:1) { # backward loop
    local_timevars$V[local_timevars$t == d] <- 
      local_timevars$u[local_timevars$t == d] + 
      pmax(local_timevars$eps_k[local_timevars$t == d] + 1 * local_timevars$EV[local_timevars$t == d + 1],
           local_timevars$eps_d[local_timevars$t == d])
  }
  
  # Calculate cumulative value
  local_timevars <- local_timevars %>% mutate(cumV = cumsum(V))
  day1 <- local_timevars %>% filter(t == 1) %>% select(cumV) %>% as.numeric()
  return(day1)
}

# solve the model for each individual with different shocks in parallel
 pred_moments <- unique(simdata$id) %>% future_map_dbl(
  ~valueday1(.x, simdata, timevars, omega, myscale=sp_treated[7]),
  .options = furrr_options(seed = TRUE)
)
pred_moments <- tibble(id = unique(simdata$id), value_day1 = pred_moments)

# now we want to shift these up and down so that extensive margin moment matches 
# TODO: do we want to include this as a parameter in the GMM estimation above? Can include for the next run
shift <- quantile(pred_moments$value_day1,1-0.2485) #quantile(pred_moments$value_day1,observed_moments$frac_observed_treated[1])
pred_moments$value_day1 <- pred_moments$value_day1 - shift
sp_treated <- c(sp_treated, shift)
```

## Counterfactuals
-    For each counterfactual, determine % using SNF, average LOS, average cost, and CS
-    Do this for healthy and sick spouse groups separately
-    Class 1: full coinsurance, no coinsurance, 20% and 50% coinsurance
-    Casss 2: coinsurance starts on day 0, 10, 21, and 50

```{r counterfactuals}
simulate_person_cf <- function(p, simdata, timevars, omega_c, omega_t,
                               myscale_c, myscale_t, shift_c, shift_t, treated=FALSE) { 
  
  if (treated == TRUE) {
    # Local copy with that person's shocks and inflate by the scale
    local_timevars <- simdata %>% filter(id == p) %>% select(eps_d, eps_k, t) %>% 
      left_join(timevars, by = "t") %>% 
      mutate(eps_d = myscale_t*eps_d, 
             eps_k = myscale_t*eps_k, # transform shocks accordingly
             EV = ifelse(t == 100, omega_t, NA))
    
    for (d in 99:1) { # backward loop
      local_timevars$EV[local_timevars$t == d] <- 
        local_timevars$u_t[local_timevars$t == d + 1] + 
        log(exp(1 * local_timevars$EV[local_timevars$t == d + 1]) + 1)
    }
    local_timevars$V <- NA
    local_timevars$V[local_timevars$t == 100] <- local_timevars$u_t[local_timevars$t == 100]
    
    for (d in 99:1) { # backward loop
      local_timevars$V[local_timevars$t == d] <- 
        local_timevars$u_t[local_timevars$t == d] + 
        pmax(local_timevars$eps_k[local_timevars$t == d] + 1 * local_timevars$EV[local_timevars$t == d + 1],
             local_timevars$eps_d[local_timevars$t == d])
    }
    
    # Calculate cumulative value
    local_timevars <- local_timevars %>% mutate(cumV = cumsum(V) - shift_t)
    
    # Find the day that maximizes cumulative value
    day1 <- local_timevars %>% filter(t == 1) %>% select(cumV) %>% as.integer()
    if (day1 >= 0) { 
      exit_day <- which.max(local_timevars$cumV)
      value <- local_timevars$cumV[exit_day]
    } else { 
      exit_day <- 0 # if negative, then exit on day 0 (never enter) 
      value <- 0
    }
    
    # If value on the first day is negative, then return 0 instead
    # exit_day <- ifelse(local_timevars$V[1] < 0, 0, exit_day)
    
    # If value is infinite, return corner solution
    exit_day <- ifelse(local_timevars$V[1] == Inf, 100, exit_day)
    return(c(exit_day,value))
  } else if (treated == FALSE) { 
     # Local copy with that person's shocks and inflate by the scale
    local_timevars <- simdata %>% filter(id == p) %>% select(eps_d, eps_k, t) %>% 
      left_join(timevars, by = "t") %>% 
      mutate(eps_d = myscale_c*eps_d, 
             eps_k = myscale_c*eps_k, # transform shocks accordingly
             EV = ifelse(t == 100, omega_c, NA))
    
    for (d in 99:1) { # backward loop
      local_timevars$EV[local_timevars$t == d] <- 
        local_timevars$u_c[local_timevars$t == d + 1] + 
        log(exp(1 * local_timevars$EV[local_timevars$t == d + 1]) + 1)
    }
    local_timevars$V <- NA
    local_timevars$V[local_timevars$t == 100] <- local_timevars$u_c[local_timevars$t == 100]
    
    for (d in 99:1) { # backward loop
      local_timevars$V[local_timevars$t == d] <- 
        local_timevars$u_c[local_timevars$t == d] + 
        pmax(local_timevars$eps_k[local_timevars$t == d] + 1 * local_timevars$EV[local_timevars$t == d + 1],
             local_timevars$eps_d[local_timevars$t == d])
    }
    
    # Calculate cumulative value
    local_timevars <- local_timevars %>% mutate(cumV = cumsum(V) - shift_c)
    
    # Find the day that maximizes cumulative value
    day1 <- local_timevars %>% filter(t == 1) %>% select(cumV) %>% as.integer()
    if (day1 >= 0) { 
      exit_day <- which.max(local_timevars$cumV)
      value <- local_timevars$cumV[exit_day]
    } else { 
      exit_day <- 0 # if negative, then exit on day 0 (never enter)  
      value <- 0
    }
    
    # If value on the first day is negative, then return 0 instead
    # exit_day <- ifelse(local_timevars$V[1] < 0, 0, exit_day)
    
    # If value is infinite, return corner solution
    exit_day <- ifelse(local_timevars$V[1] == Inf, 100, exit_day)
    return(c(exit_day,value))
    }
}

N <- 10000
simdata <- expand_grid(
  id = 1:N,
  t = 1:100
) %>%
  arrange(id, t) %>% 
  mutate(eps_k = revd(n = (100*N), scale=1), 
         eps_d = revd(n = (100*N), scale=1)) 

l0_c <- sp_control[1] #scaling
l1_c <- sp_control[2]/1000 #scaling
l2_c <- sp_control[3]/1000 #scaling
l3_c <- sp_control[4]/1000 # scaling
alpha_c <- sp_control[5]
omega_c <- sp_control[6]
shift_c <- sp_control[8]

l0_t <- sp_treated[1] #scaling
l1_t <- sp_treated[2]/1000 #scaling
l2_t <- sp_treated[3]/1000 #scaling
l3_t <- sp_treated[4]/1000 # scaling
alpha_t <- sp_treated[5]
omega_t <- sp_treated[6]
shift_t <- sp_treated[8]

# make dataframe 
timevars <- data.frame(t = 1:100) %>%
  mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
         lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
  mutate(p = ifelse(t < 22, 1e-5, 37.5),
         omega_c = omega_c,
         omega_t = omega_t,
         delta = 1,
         alpha_c = alpha_c,
         alpha_t = alpha_t,
         u_c = lambda_c + alpha_c*p, 
         u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
cfsoln_t <- unique(simdata$id) %>% future_map(
  ~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                      myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                      shift_c, shift_t, treated=TRUE),
  .options = furrr_options(seed = TRUE)
)
cfsoln_t <- as.data.frame(do.call(rbind, cfsoln_t))
names(cfsoln_t) <- c("exit_day", "value")
 cfsoln_c <- unique(simdata$id) %>% future_map(
  ~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                      myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                      shift_c, shift_t, treated=FALSE),
  .options = furrr_options(seed = TRUE)
)
 cfsoln_c <- as.data.frame(do.call(rbind, cfsoln_c))
 names(cfsoln_c) <- c("exit_day", "value")

# calculate main table measures
cfsoln_t %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_t %>% summarize(exit_day = mean(exit_day))
cfsoln_t %>% mutate(paid = pmax(0,(exit_day-20)*37.5)) %>% 
  summarize(paid = mean(paid))
cfsoln_t %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0.8)))) %>% 
  summarize(paid = mean(paid)) # government costs: full coverage for the first days, then 80% of the rest (assuming no Medigap for now). 
cfsoln_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value))

cfsoln_c %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_c %>% summarize(exit_day = mean(exit_day))
cfsoln_c %>% 
  mutate(paid = pmax(0,(exit_day-20)*37.5)) %>% 
  summarize(paid = mean(paid))
cfsoln_c %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0.8)))) %>% 
  summarize(paid = mean(paid)) # government costs: full coverage for the first days, then 80% of the rest (assuming no Medigap for now). 
cfsoln_c %>% 
  mutate(value = value/0.097) %>% summarize(value = mean(value))
```

Now repeat this for price at 0%, 50%, and 100% coinsurance 

```{r cf-price}
# first, set price to 0 
timevars <- data.frame(t = 1:100) %>%
  mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
         lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
  mutate(p = ifelse(t < 22, 1e-5, 1e-5),
         omega_c = omega_c,
         omega_t = omega_t,
         delta = 1,
         alpha_c = alpha_c,
         alpha_t = alpha_t,
         u_c = lambda_c + alpha_c*p, 
         u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
cfsoln_p0_t <- unique(simdata$id) %>% future_map(
  ~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                      myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                      shift_c, shift_t, treated=TRUE),
  .options = furrr_options(seed = TRUE)
)
cfsoln_p0_t <- as.data.frame(do.call(rbind, cfsoln_p0_t))
names(cfsoln_p0_t) <- c("exit_day", "value")
 cfsoln_p0_c <- unique(simdata$id) %>% future_map(
  ~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                      myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                      shift_c, shift_t, treated=FALSE),
  .options = furrr_options(seed = TRUE)
)
 cfsoln_p0_c <- as.data.frame(do.call(rbind, cfsoln_p0_c))
 names(cfsoln_p0_c) <- c("exit_day", "value")
 
 cfsoln_p0_t %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_p0_t %>% summarize(exit_day = mean(exit_day))
cfsoln_p0_t %>% mutate(paid = pmax(0,(exit_day-20)*187.5*0)) %>% 
  summarize(paid = mean(paid))
cfsoln_p0_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value))
cfsoln_p0_t %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*1)))) %>% 
  summarize(paid = mean(paid)) # government costs: 

cfsoln_p0_c %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_p0_c %>% summarize(exit_day = mean(exit_day))
cfsoln_p0_c %>% mutate(paid = pmax(0,(exit_day-20)*187.50*0)) %>% 
  summarize(paid = mean(paid))
cfsoln_p0_c %>% mutate(value = value/0.097) %>% summarize(value = mean(value))
cfsoln_p0_c %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*1)))) %>% 
  summarize(paid = mean(paid)) # government costs: 
 
# now 50% coinsurance
timevars <- data.frame(t = 1:100) %>%
mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
       lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
mutate(p = ifelse(t < 22, 1e-5, 187.50*0.5),
       omega_c = omega_c,
       omega_t = omega_t,
       delta = 1,
       alpha_c = alpha_c,
       alpha_t = alpha_t,
       u_c = lambda_c + alpha_c*p, 
       u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
cfsoln_p50_t <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=TRUE),
.options = furrr_options(seed = TRUE)
)
cfsoln_p50_t <- as.data.frame(do.call(rbind, cfsoln_p50_t))
names(cfsoln_p50_t) <- c("exit_day", "value")
cfsoln_p50_c <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=FALSE),
.options = furrr_options(seed = TRUE)
)
cfsoln_p50_c <- as.data.frame(do.call(rbind, cfsoln_p50_c))
names(cfsoln_p50_c) <- c("exit_day", "value")

 cfsoln_p50_t %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_p50_t %>% summarize(exit_day = mean(exit_day))
cfsoln_p50_t %>% mutate(paid = pmax(0,(exit_day-20)*187.5*0.5)) %>% 
  summarize(paid = mean(paid))
cfsoln_p50_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value))
cfsoln_p50_t %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0.5)))) %>% 
  summarize(paid = mean(paid)) # government costs: 

cfsoln_p50_c %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_p50_c %>% summarize(exit_day = mean(exit_day))
cfsoln_p50_c %>%  mutate(paid = pmax(0,(exit_day-20)*187.50*0.5)) %>% 
  summarize(paid = mean(paid))
cfsoln_p50_c %>%  mutate(value = value/0.097) %>% summarize(value = mean(value))
cfsoln_p50_c %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0.5)))) %>% 
  summarize(paid = mean(paid)) # government costs: 

# now 100% coinsurance
timevars <- data.frame(t = 1:100) %>%
mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
       lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
mutate(p = ifelse(t < 22, 1e-5, 187.50*1),
       omega_c = omega_c,
       omega_t = omega_t,
       delta = 1,
       alpha_c = alpha_c,
       alpha_t = alpha_t,
       u_c = lambda_c + alpha_c*p, 
       u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
cfsoln_p100_t <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=TRUE),
.options = furrr_options(seed = TRUE)
)
cfsoln_p100_t <- as.data.frame(do.call(rbind, cfsoln_p100_t))
names(cfsoln_p100_t) <- c("exit_day", "value")
cfsoln_p100_c <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=FALSE),
.options = furrr_options(seed = TRUE)
)
cfsoln_p100_c <- as.data.frame(do.call(rbind, cfsoln_p100_c))
names(cfsoln_p100_c) <- c("exit_day", "value")

cfsoln_p100_t %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_p100_t %>% summarize(exit_day = mean(exit_day))
cfsoln_p100_t %>% mutate(paid = pmax(0,(exit_day-20)*187.5*0.5)) %>% 
  summarize(paid = mean(paid))
cfsoln_p100_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value))
cfsoln_p100_t %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0)))) %>% 
  summarize(paid = mean(paid)) # government costs: 

cfsoln_p100_c %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_p100_c %>% summarize(exit_day = mean(exit_day))
cfsoln_p100_c %>% mutate(paid = pmax(0,(exit_day-20)*187.50*1)) %>% 
  summarize(paid = mean(paid))
cfsoln_p100_c %>% mutate(value = value/0.097) %>% summarize(value = mean(value))
cfsoln_p100_c %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0)))) %>% 
  summarize(paid = mean(paid)) # government costs: 
```

Now vary when that coinsurance starts
```{r cf-time}

# day 0
timevars <- data.frame(t = 1:100) %>%
mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
       lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
mutate(p = 187.50*0.2,
       omega_c = omega_c,
       omega_t = omega_t,
       delta = 1,
       alpha_c = alpha_c,
       alpha_t = alpha_t,
       u_c = lambda_c + alpha_c*p, 
       u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
cfsoln_d0_t <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=TRUE),
.options = furrr_options(seed = TRUE)
)
cfsoln_d0_t <- as.data.frame(do.call(rbind, cfsoln_d0_t))
names(cfsoln_d0_t) <- c("exit_day", "value")
cfsoln_d0_c <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=FALSE),
.options = furrr_options(seed = TRUE)
)
cfsoln_d0_c <- as.data.frame(do.call(rbind, cfsoln_d0_c))
names(cfsoln_d0_c) <- c("exit_day", "value")

 cfsoln_d0_t %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_d0_t %>% summarize(exit_day = mean(exit_day))
cfsoln_d0_t %>% mutate(paid = pmax(0,(exit_day-0)*187.5*0.2)) %>%
  summarize(paid = mean(paid))
cfsoln_d0_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value))

cfsoln_d0_c %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_d0_c %>% summarize(exit_day = mean(exit_day))
cfsoln_d0_c %>% mutate(paid = pmax(0,(exit_day-0)*187.50*0.2)) %>%
  summarize(paid = mean(paid))
cfsoln_d0_c %>% mutate(value = value/0.097) %>% summarize(value = mean(value))

# day 7
timevars <- data.frame(t = 1:100) %>%
mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
       lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
mutate(p = ifelse(t < 8, 1e-5, 187.50*0.2),
       omega_c = omega_c,
       omega_t = omega_t,
       delta = 1,
       alpha_c = alpha_c,
       alpha_t = alpha_t,
       u_c = lambda_c + alpha_c*p, 
       u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
cfsoln_d7_t <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=TRUE),
.options = furrr_options(seed = TRUE)
)
cfsoln_d7_t <- as.data.frame(do.call(rbind, cfsoln_d7_t))
names(cfsoln_d7_t) <- c("exit_day", "value")
cfsoln_d7_c <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=FALSE),
.options = furrr_options(seed = TRUE)
)
cfsoln_d7_c <- as.data.frame(do.call(rbind, cfsoln_d7_c))
names(cfsoln_d7_c) <- c("exit_day", "value")

 cfsoln_d7_t %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_d7_t %>% summarize(exit_day = mean(exit_day))
cfsoln_d7_t %>% mutate(paid = pmax(0,(exit_day-7)*187.5*0.2)) %>%
  summarize(paid = mean(paid))
cfsoln_d7_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value))

cfsoln_d7_c %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_d7_c %>%  summarize(exit_day = mean(exit_day))
cfsoln_d7_c %>% mutate(paid = pmax(0,(exit_day-7)*187.50*0.2)) %>%
  summarize(paid = mean(paid))
cfsoln_d7_c %>% mutate(value = value/0.097) %>% summarize(value = mean(value))

# day 30
timevars <- data.frame(t = 1:100) %>%
mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
       lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
mutate(p = ifelse(t < 31, 1e-5, 187.50*0.2),
       omega_c = omega_c,
       omega_t = omega_t,
       delta = 1,
       alpha_c = alpha_c,
       alpha_t = alpha_t,
       u_c = lambda_c + alpha_c*p, 
       u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
cfsoln_d30_t <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=TRUE),
.options = furrr_options(seed = TRUE)
)
cfsoln_d30_t <- as.data.frame(do.call(rbind, cfsoln_d30_t))
names(cfsoln_d30_t) <- c("exit_day", "value")
cfsoln_d30_c <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=FALSE),
.options = furrr_options(seed = TRUE)
)
cfsoln_d30_c <- as.data.frame(do.call(rbind, cfsoln_d30_c))
names(cfsoln_d30_c) <- c("exit_day", "value")

 cfsoln_d30_t %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_d30_t %>% summarize(exit_day = mean(exit_day))
cfsoln_d30_t %>% mutate(paid = pmax(0,(exit_day-30)*187.5*0.2)) %>%
  summarize(paid = mean(paid))
cfsoln_d30_t %>%  mutate(value = value/0.009) %>% summarize(value = mean(value))

cfsoln_d30_c %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_d30_c %>% summarize(exit_day = mean(exit_day))
cfsoln_d30_c %>% mutate(paid = pmax(0,(exit_day-30)*187.50*0.2)) %>%
  summarize(paid = mean(paid))
cfsoln_d30_c %>%  mutate(value = value/0.097) %>% summarize(value = mean(value))

# day 60
timevars <- data.frame(t = 1:100) %>%
mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
       lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
mutate(p = ifelse(t < 61, 1e-5, 187.50*0.2),
       omega_c = omega_c,
       omega_t = omega_t,
       delta = 1,
       alpha_c = alpha_c,
       alpha_t = alpha_t,
       u_c = lambda_c + alpha_c*p, 
       u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
cfsoln_d60_t <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=TRUE),
.options = furrr_options(seed = TRUE)
)
cfsoln_d60_t <- as.data.frame(do.call(rbind, cfsoln_d60_t))
names(cfsoln_d60_t) <- c("exit_day", "value")
cfsoln_d60_c <- unique(simdata$id) %>% future_map(
~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                    myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                    shift_c, shift_t, treated=FALSE),
.options = furrr_options(seed = TRUE)
)
cfsoln_d60_c <- as.data.frame(do.call(rbind, cfsoln_d60_c))
names(cfsoln_d60_c) <- c("exit_day", "value")

 cfsoln_d60_t %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_d60_t %>% summarize(exit_day = mean(exit_day))
cfsoln_d60_t %>% mutate(paid = pmax(0,(exit_day-60)*187.5*0.2)) %>%
  summarize(paid = mean(paid))
cfsoln_d60_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value))

cfsoln_d60_c %>% mutate(test = ifelse(exit_day > 0, 100, 0)) %>% summarize(test = mean(test))
cfsoln_d60_c %>% summarize(exit_day = mean(exit_day))
cfsoln_d60_c %>% mutate(paid = pmax(0,(exit_day-60)*187.50*0.2)) %>%
  summarize(paid = mean(paid))
cfsoln_d60_c %>% mutate(value = value/0.097) %>% summarize(value = mean(value))

beepr::beep(sound=3)
```

## Transfers with Balanced Budget

```{r transfers}
# per-person government spending prior to any transfers (weighted by popsize)
frac_c <- 1-0.27 # from KFF, about 27% of Medicare enrollees live alone
pre_spend <- (1-frac_c)*(cfsoln_t %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0.8)))) %>% 
  summarize(paid = mean(paid)) %>% pull(paid)) + frac_c*(cfsoln_c %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0.8)))) %>% 
  summarize(paid = mean(paid)) %>% pull(paid))

# if we move to full insurance for the treated group, what is per-person government spending only in that group
post_spend_t <- cfsoln_p0_t %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*1)))) %>% 
  summarize(paid = mean(paid)) %>% pull(paid)

# What does this mean, for a balanced budget, can be spent per person on the control group? 
post_spend_c <- (pre_spend - (1-frac_c)*post_spend_t)/frac_c

# What does this level of spending imply for the level of coverage that can be offered to the control group? (rough estimate suggests that coverage should be about 71.2% of what it was, or 56.96% instead of 80%)
timevars <- data.frame(t = 1:100) %>%
  mutate(lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3, 
         lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
  mutate(p = ifelse(t < 22, 1e-5, 187.50*(1-0.8*0.95)), # new price is $45 instead of $37.50
         omega_c = omega_c,
         omega_t = omega_t,
         delta = 1,
         alpha_c = alpha_c,
         alpha_t = alpha_t,
         u_c = lambda_c + alpha_c*p, 
         u_t = lambda_t + alpha_t*p) 

# solve the model for each individual with different shocks in parallel
 cfsoln_transfer_c <- unique(simdata$id) %>% future_map(
  ~simulate_person_cf(.x, simdata, timevars, omega_c, omega_t, 
                      myscale_c=sp_control[7], myscale_t = sp_treated[7], 
                      shift_c, shift_t, treated=FALSE),
  .options = furrr_options(seed = TRUE)
)
cfsoln_transfer_c <- as.data.frame(do.call(rbind, cfsoln_transfer_c))
names(cfsoln_transfer_c) <- c("exit_day", "value")
post_spend <- (1-frac_c)*(cfsoln_p0_t %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*1)))) %>% 
  summarize(paid = mean(paid)) %>% pull(paid)) + frac_c*(cfsoln_transfer_c %>% mutate(paid = ifelse(exit_day == 0, 0, 
                                  ifelse(exit_day <= 20, exit_day*187.50, 
                                  20*187.50 + (exit_day-20)*(187.50*0.8*0.712)))) %>% 
  summarize(paid = mean(paid)) %>% pull(paid))

pre_welfare <- frac_c*(cfsoln_c %>% mutate(value = value/0.097) %>% summarize(value = mean(value)) %>% pull(value)) + 
  (1-frac_c)*(cfsoln_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value)) %>% pull(value))
post_welfare <- frac_c*(cfsoln_transfer_c %>% mutate(value = value/0.097) %>% summarize(value = mean(value)) %>% pull(value)) + 
  (1-frac_c)*(cfsoln_p0_t %>% mutate(value = value/0.009) %>% summarize(value = mean(value)) %>% pull(value))
```

## Main Equations

-   utility flows: $u_t = \lambda_t + \alpha p_t$ (where we use the SNF price schedule for $p_t$)
-   patient decision: $\{$enter,exit$\}$ with idiosyncratic (T1EV) shocks $\{\varepsilon_{kt}, \varepsilon_{dt}\}$
-   Hence $V_t(\varepsilon_t) = u_t + \max\{\varepsilon_{kt}+\delta EV_{t+1}, \varepsilon_{dt}\}$
-   Assign a termination value $\Omega=EV_{100}$
-   Given distributional assumptions, this gives us $EV_t=u_{t+1}+\log(\exp(\delta EV_{t+1})+1)$

## Parameterization

```{r parameterization }
# Load estimated parameters from both estimations 
rm(list = ls())
load(paste0(path, "Convergence_DynamicModel_TREATED_20250410.RData", sep=""))
sp_treated <- myouts$coefficients
rm(list = setdiff(ls(), c("path", "sp_treated")))
load(paste0(path, "Convergence_DynamicModel_CONTROL_20250410.RData", sep=""))
sp_control <- myouts$coefficients
tokeep <- c("sp_treated", "sp_control", "moment_function", "simulate_person", "simulate_moments")
rm(list = setdiff(ls(), tokeep))

# structparams[1] -- structparams[4] are the parameters for lambda_t 
# structparams[5] is alpha (should be negative) 
# structparams[6] is omega (terminal value of SNF at day 100) 
# structparams[7] governs the variance of the shocks 
# structparams[8]-structparams[10] for day-specific utility shocks for day 21, 22, 23

# Do this once for treated and once for control 
l0_t <- sp_treated[1] # these variables dictate $\lambda_t$
l1_t <- sp_treated[2]/1000
l2_t <- sp_treated[3]/1000
l3_t <- sp_treated[4]/1000
omega_t <- sp_treated[6] # termination value
delta_t <- 1 # discount factor
alpha_t <- sp_treated[5] # price coefficient (for treated group; higher for control group)
epscale_t <- sp_treated[7] # scaling for shocks

l0_c <- sp_control[1] # these variables dictate $\lambda_c$
l1_c <- sp_control[2]/1000
l2_c <- sp_control[3]/1000
l3_c <- sp_control[4]/1000
omega_c <- sp_control[6] # termination value
delta_c <- 1 # discount factor
alpha_c <- sp_control[5] # price coefficient (for treated group; higher for control group)
epscale_c <- sp_control[7] # scaling for shocks

# make a dataframe with 100 periods and a variable t
timevars <- data.frame(t = 1:100) %>% 
  mutate(lambda_t = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3, 
         lambda_c = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3) %>%
  # mutate(ubump_t = ifelse(t == 21, sp_treated[8], 
  #                         ifelse(t == 22, sp_treated[9], 
  #                                ifelse(t == 23, sp_treated[10], 0))),
  #        ubump_c = ifelse(t == 21, sp_control[8], 
  #                         ifelse(t == 22, sp_control[9], 
  #                                ifelse(t == 23, sp_control[10], 0)))) %>%
  mutate(p = ifelse(t < 22, 1e-5, 37.5), 
         omega_t = omega_t, 
         delta = delta_t,
         alpha_t = alpha_t,
         omega_c = omega_c, 
         alpha_c = alpha_c,
         u_t = lambda_t + alpha_t*p, # + ubump_t,
         u_c = lambda_c + alpha_c*p) # + ubump_c)

# Finally, draw the shocks in each period using a T1EV distribution
timevars <- timevars %>% 
  mutate(eps_k = revd(n = 100, scale=3), 
         eps_d = revd(n = 100, scale=3))

# Plot the utility curves for both groups
ggplot(timevars, aes(x=t)) + 
  geom_point(aes(y=u_t, color='Sick Spouse Group'), size=2) +
  geom_point(aes(y=u_c, color='Healthy Spouse Group'), size=2) +
  theme_minimal() + 
  labs(x = "Days in SNF", y = "", subtitle="Flow Utilities", color=NULL) + 
  geom_hline(yintercept = 0) # + 
  ggplot::theme(legend.position="inside",
      legend.position.inside = c(0.01, 0.99),       # top-left corner inside the plot area
      legend.justification = c("left", "top"))
```

## Calculate EV via the Bellman equation
    - Given this, we then use the shocks to calculate the value function in each period and maximize
    - We then calculate and output cumulative value, choose to exit to maximize this 

```{r bellman, }
timevars <- timevars %>% mutate(EV_t = ifelse(t==100, omega_t, NA),
                                EV_c = ifelse(t==100, omega_c, NA))
for (d in 99:1) { # backward loop
  timevars <- timevars %>% 
    mutate(EV_t = ifelse(t == d, lead(u_t) + log(exp(delta*lead(EV_t))+1), EV_t),
           EV_c = ifelse(t == d, lead(u_c) + log(exp(delta*lead(EV_c))+1), EV_c))
}

ggplot(timevars, aes(x=t)) + 
  geom_point(aes(y=EV_t, color='Sick Spouse Group'), size=2) +
  geom_point(aes(y=EV_c, color='Healthy Spouse Group'), size=2) +
  theme_minimal() + 
  labs(x = "Days in SNF", y = "", subtitle="Expected Value at t", color=NULL) + 
  geom_hline(yintercept = 0) # + 
  theme(legend.position="inside",
      legend.position.inside = c(0.01, 0.99),       # top-left corner inside the plot area
      legend.justification = c("left", "top"))
```

## Extrapolating Demand
    - Given these expected valuations, exit day is day that maximizes expected cumulative value, with individual heterogeneity
    - We can then pull demand for day 21 as the # of individuals in sample choosing to stay as price varies
```{r demand}
plan(multisession, workers = parallel::detectCores() - 1)

# Parameters to specify
N <- 1000 # how many individuals to sample
myday <- 22 # which day are you constructing the demand curve for? 

# Functions (3)
simulate_person <- function(p, simdata, timevars, omega, myscale) { 
    #  A needed function to use a given person's shocks to return outcomes
  # Local copy with that person's shocks and inflate by the scale
  local_timevars <- simdata %>% filter(id == p) %>% select(eps_d, eps_k, t) %>% 
    left_join(timevars, by = "t") %>% 
    mutate(eps_d = myscale*eps_d, 
           eps_k = myscale*eps_k, # transform shocks accordingly
           EV = ifelse(t == 100, omega, NA))
  
  for (d in 99:1) { # backward loop
    local_timevars$EV[local_timevars$t == d] <- 
      local_timevars$u[local_timevars$t == d + 1] + 
      log(exp(1 * local_timevars$EV[local_timevars$t == d + 1]) + 1)
  }
  local_timevars$V <- NA
  local_timevars$V[local_timevars$t == 100] <- local_timevars$u[local_timevars$t == 100]
  
  for (d in 99:1) { # backward loop
    local_timevars$V[local_timevars$t == d] <- 
      local_timevars$u[local_timevars$t == d] + 
      pmax(local_timevars$eps_k[local_timevars$t == d] + 1 * local_timevars$EV[local_timevars$t == d + 1],
           local_timevars$eps_d[local_timevars$t == d])
  }
  
  # Calculate cumulative value
  local_timevars <- local_timevars %>% mutate(cumV = cumsum(V))
  
  # Find the day that maximizes cumulative value
  exit_day <- which.max(local_timevars$cumV)
  
  # If value on the first day is negative, then return 0 instead
  # exit_day <- ifelse(local_timevars$V[1] < 0, 0, exit_day)
  
  # If value is infinite, return corner solution
  exit_day <- ifelse(local_timevars$V[1] == Inf, 100, exit_day)
  return(exit_day)
}

return_demand_c <- function(myprice) {
# FIRST DO THIS FOR THE CONTROL (HEALTHY SPOUSE) GROUP
  # make a dataframe with 100 periods and a variable t
  timevars <- data.frame(t = 1:100) %>%
    mutate(lambda = l0_c + l1_c*t + l2_c*t^2 + l3_c*t^3) %>%
    mutate(p = ifelse(t < 22, 1e-5, myprice),
           omega = omega_c,
           delta = 1,
           alpha = alpha_c,
           u = lambda + alpha*p ) 
  
  # solve the model for each individual with different shocks in parallel
   pred_moments <- unique(simdata$id) %>% future_map_int(
    ~simulate_person(.x, simdata, timevars, omega, myscale=epscale_c),
    .options = furrr_options(seed = TRUE)
  )
  pred_moments_c <- tibble(id = unique(simdata$id), exit_day = pred_moments) %>% 
    mutate(insnf = (exit_day >= myday)) %>% summarize(mean = mean(insnf)) %>% pull(mean)
  pred_moments_c
} 

return_demand_t <- function(myprice) {
# SECOND, DO THIS FOR THE TREATMENT (SICK SPOUSE) GROUP
  # make a dataframe with 100 periods and a variable t
  timevars <- data.frame(t = 1:100) %>%
    mutate(lambda = l0_t + l1_t*t + l2_t*t^2 + l3_t*t^3) %>%
    mutate(p = ifelse(t < 22, 1e-5, myprice),
           omega = omega_t,
           delta = 1,
           alpha = alpha_t,
           u = lambda + alpha*p ) 
  
  # solve the model for each individual with different shocks in parallel
   pred_moments <- unique(simdata$id) %>% future_map_int(
    ~simulate_person(.x, simdata, timevars, omega, myscale=epscale_t),
    .options = furrr_options(seed = TRUE)
  )
  pred_moments_t <- tibble(id = unique(simdata$id), exit_day = pred_moments) %>% 
     mutate(insnf = (exit_day >= myday)) %>% summarize(mean = mean(insnf)) %>% pull(mean)
  pred_moments_t
}

# make a dataset of N individuals with unique shocks (drawn once for all iterations)
simdata <- expand_grid(
  id = 1:N,
  t = 1:100
) %>%
  arrange(id, t) %>% 
  mutate(eps_k = revd(n = (100*N), scale=1), 
         eps_d = revd(n = (100*N), scale=1)) 

# Vary across prices
prices <- seq(from=0,to=2500,by=100)
demand_c <- map_dbl(prices, return_demand_c)
demand_t <- map_dbl(prices, return_demand_t)

# Graph 
tibble(
  price = prices,
  HealthySpouse = demand_c,
  SickSpouse = demand_t
) %>% 
mutate(SickSpouse = SickSpouse * 1.191) %>%  # Scale by extensive margin differences across groups 
  pivot_longer(cols = c(HealthySpouse, SickSpouse), 
               names_to = "Group", 
               values_to = "Demand") %>% 
  ggplot(aes(x = Demand, y = price, color = Group)) +
    geom_line(linewidth = 1.2) +
    geom_point() +
    theme_minimal() +
    labs(
      title = "Estimated Demand Curves: Day 21",
      x = "Q",
      y = "P",
      color = "Group"
    )
```

## Consumer Surplus
    - Repeat demand analysis above for days 0 through 100 
    - Estimate CS as area under the demand curve for each day 
```{r cs}
simulate_person <- function(p, simdata, timevars, omega, myscale) {
  # Extract this personŌĆÖs shocks from simdata and join with timevars.
  local_timevars <- simdata %>%
    filter(id == p) %>%
    select(eps_d, eps_k, t) %>%
    left_join(timevars, by = "t") %>%
    # Scale the shocks accordingly.
    mutate(eps_d = myscale * eps_d, 
           eps_k = myscale * eps_k)
  
  # We assume the joined data is sorted in ascending order by t.
  # Extract the required vectors.
  u <- local_timevars$u
  eps_d <- local_timevars$eps_d
  eps_k <- local_timevars$eps_k
  
  # Call the Rcpp function which performs the heavy computations.
  simulate_person_cpp(u, eps_d, eps_k, omega)
}


# Update functions to loop over days
# rm(myprice, u)
simulate_demand <- function(myprice,myday,
                            l0,l1,l2,l3,omega_val,alpha_val,epscale_val) {
  # make a dataframe with 100 periods and a variable t
  timevars <- data.frame(t = 1:100) %>%
    mutate(lambda = l0 + l1*t + l2*t^2 + l3*t^3) %>%
    mutate(p = ifelse(t < 22, 1e-5, myprice),
           omega = omega_val,
           delta = 1,
           alpha = alpha_val,
           u = lambda + alpha*p ) 
  
  # solve the model for each individual with different shocks in parallel
   pred_moments <- ids %>% map_int(
    ~simulate_person(.x, simdata, timevars, omega_val, myscale=epscale_val))
  tibble(id = ids, exit_day = pred_moments) %>% 
    mutate(insnf = (exit_day >= myday)) %>% summarize(mean = mean(insnf)) %>% pull(mean)
} 

return_demand_c <- function(myprice, myday) { 
  simulate_demand(myprice,myday, 
                  l0_c,l1_c,l2_c,l3_c,
                  omega_c,alpha_c, epscale_c)
}
return_demand_t <- function(myprice, myday) { 
  simulate_demand(myprice,myday, 
                  l0_t,l1_t,l2_t,l3_t,
                  omega_t,alpha_t, epscale_t)
  }

# Create an explicit PSOCK cluster using available cores (minus one)
cl <- makeCluster(detectCores() - 1)
plan(cluster, workers=cl) 

# Update simulate_person to use Rcpp to reduce bottleneck
parallel::clusterEvalQ(cl, {
  library(Rcpp)
  library(tidyverse)
  sourceCpp(paste0(path, "simulate_person.cpp", sep="")) 
})
clusterExport(cl, varlist = c("simulate_person", "simulate_demand", 
                              "return_demand_c", "return_demand_t", 
                              "simdata", "ids",
                              # Export any constants or additional variables, e.g.:
                              "l0_c", "l1_c", "l2_c", "l3_c", "omega_c", "alpha_c", "epscale_c",
                              "l0_t", "l1_t", "l2_t", "l3_t", "omega_t", "alpha_t", "epscale_t"))

# use the set of shocks drawn above
ids <- unique(simdata$id)

# Vary across prices *and days*
output <- expand_grid(price = seq(from=0,to=2500,by=50),  
                      day = 1:50) # TODO: extend to 100 days

# Convert the grid into a list of rows
grid_list <- split(output, seq(nrow(output)))

# Use parLapply to loop over each row
results <- parLapply(cl, grid_list, function(row) {
  price <- row$price
  day <- row$day
  
  # Call your demand functions
  demand_c <- return_demand_c(price, day) * 0.01957
  demand_t <- return_demand_t(price, day) * 0.02457
    # Adjust for extensive margin responses in demand curves 
  
  # Return a small data.frame for this parameter set
  data.frame(price = price, day = day, demand_c = demand_c, demand_t = demand_t)
})

# Combine the results into one data.frame
final_results <- do.call(rbind, results)
output <- final_results
rm(final_results)

# output <- output %>%
#   mutate(
#     demand_c = future_map2_dbl(price, day, ~ return_demand_c(.x, .y)),
#     demand_t = future_map2_dbl(price, day, ~ return_demand_t(.x, .y))
#   ) %>% mutate(demand_t = demand_t * 1.191) 

# here's the nice version of the demand curve
demandfig <- output %>% 
    filter(day==22) %>% filter(price < 500) %>% 
    filter(demand_c > 0 & demand_t > 0) %>% 
    pivot_longer(cols = c(demand_c, demand_t),
                 names_to = "Group",
                 values_to = "Demand") %>% 
    arrange(Demand) %>%
    ggplot(aes(x = Demand*100, y = price, color = Group)) +
    geom_line(linewidth = 1.2) +
    geom_point() +
    theme_minimal() +
    labs(
        title = paste0("Estimated Demand Curves: Day ", 22, sep=""),
        x = "% of Outcome Spouses Demanding SNF Care",
        y = "P",
        color = "Group"
    ) + geom_hline(yintercept = 37.5, linetype="dashed") +  
  geom_text(aes(x = 1.65, y = 38, label = "Expected Price = $37.50"),
            color = "black", vjust = -1) + 
  scale_color_discrete(labels = c("Healthy Spouses", "Sick Spouses")) +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))


# sense check: draw the demand curve again for each day
for (d in 1:50) {
demandfig <- output %>% 
  filter(day==d) %>% 
  filter(demand_c > 0 & demand_t > 0) %>% 
  pivot_longer(cols = c(demand_c, demand_t),
               names_to = "Group",
               values_to = "Demand") %>% 
  arrange(Demand) %>%
  ggplot(aes(x = Demand*100, y = price, color = Group)) +
    geom_line(linewidth = 1.2) +
    geom_point() +
    theme_minimal() +
    labs(
      title = paste0("Estimated Demand Curves: Day ", d, sep=""),
      x = "Q",
      y = "P",
      color = "Group"
    )
ggsave(demandfig, 
       file=paste0(path, "DemandCurve_Day",d,".png",
                   sep="" ), width=8, height=4, dpi=300)
}

# Calculate CS per day by integrating over prices 
  # Notes: for early days there's not a lot of demand variation -- set CS to be 0 when this happens
test <- output %>% group_by(day) %>% arrange(demand_c) %>% 
  mutate(cs_c = trapz(demand_c, price)*100,
         cs_t = trapz(demand_t, price)*100) %>% 
  select(day, cs_c, cs_t) %>% unique() # one observation per day

test %>% pivot_longer(cols = c(cs_c, cs_t), 
               names_to = "Group",
               values_to = "CS") %>% 
ggplot(aes(x=day, y=CS, color=Group)) + geom_point() + 
  geom_line(linewidth=1.2) + theme_minimal() + 
  scale_color_manual(values = c("skyblue2", "maroon3"), 
                     labels = c("Healthy Spouse", "Sick Spouse")) +
  labs(title = "Estimated CS by Day", 
       x = "Day", 
       y = "$", 
       color="Group")

# Now cumulative
test %>% ungroup() %>% arrange(day) %>% mutate(cum_c = cumsum(cs_c), 
                                  cum_t = cumsum(cs_t)) %>% 
  pivot_longer(cols = c(cum_c, cum_t), 
               names_to = "Group",
               values_to = "CS") %>% 
ggplot(aes(x=day, y=CS, color=Group)) + geom_point() + 
  geom_line(linewidth=1.2) + theme_minimal() + 
  scale_color_manual(values = c("skyblue2", "maroon3"), 
                     labels = c("Healthy Spouse", "Sick Spouse")) +
  labs(title = "Estimated Cumulative CS by Day", 
       x = "Day", 
       y = "$", 
       color="Group")

# What is ratio of cumulative CS over days?
test %>% ungroup() %>% arrange(day) %>% mutate(cum_c = cumsum(cs_c), 
                                  cum_t = cumsum(cs_t)) %>% 
  pivot_longer(cols = c(cum_c, cum_t), 
               names_to = "Group",
               values_to = "CS") %>% 
  mutate(ratio = ifelse(cs_c == 0, 1, cs_t / cs_c)) %>%
ggplot(aes(x=day, y=ratio)) + geom_point() + 
  geom_line(linewidth=1.2) + theme_minimal() + 
  geom_hline(yintercept = 3.10, linetype='dashed') + 
  labs(title = "Ratio of Treated:Control Cumulative CS by Day", 
       x = "Day", 
       y = "")
```

## Ratio of Utilities (Policy) 
What if we could think about moving either when deductibles started or the amount of coinusrance? How would this change ratio of utilities at the day when copayments start?

```{r policy-variation}
# First, let's fix the price at 37.50 and just move when it "turns on" 
# If you make this utilities on day of copayment, that doesn't change much across groups
# If you make it utilities on day 22, then it doesn't change at all except discretely
# So take cumulative expected utility from day 1 to day 
ratio_days <- function(day, treated=FALSE) { 
  if (treated == TRUE) { 
    # return(l0_t + l1_t*day + l2_t*(day)^2 + l3_t*(day)^3 + alpha_t * 37.50) # one day
    return(l0_t * day + (l1_t * day * (day + 1) / 2) + (l2_t * day * (day + 1) * (2*day + 1) / 6) + (l3_t * (day * (day + 1) / 2)^2) + alpha_t * 37.50) # cumulative u
  } else if (treated == FALSE) {
  # return(l0_c + l1_c*day + l2_c*(day)^2 + l3_c*(day)^3 + alpha_c * 37.50) # one day 
    return(l0_c * day + (l1_c * day * (day + 1) / 2) + (l2_c * day * (day + 1) * (2*day + 1) / 6) + (l3_c * (day * (day + 1) / 2)^2) + alpha_c * 37.50)# cumulative u
  } 
  # return((u_t-u_c)/abs(u_c)*100)
}

days <- seq(from=10,to=40,by=1)
ratios_c <- ratio_days(days, treated = FALSE)
ratios_t <- ratio_days(days, treated = TRUE)
data.frame(cbind(days,Treatment=ratios_t, Control=ratios_c)) %>% 
  pivot_longer(
    cols      = c(Treatment, Control),
    names_to  = "Group",
    values_to = "ratio"
  ) %>%
  ggplot(aes(x=days,y=ratio,color=Group,fill=Group)) + 
  geom_point() + 
  theme_minimal() + 
  geom_vline(xintercept = 21.5, linetype='dashed') + 
  labs(x="Day Copayments Start", y="Cumulative Utilities", fill="Group")



# Secon, let the coinusrance rate vary and keep it fixed at day 22
ratio_price <- function(price, treated=FALSE) { 
  if (treated == TRUE) { 
    return(l0_t + l1_t*22 + l2_t*(22)^2 + l3_t*(22)^3 + alpha_t * 187.5*price)
  } else if (treated == FALSE) {
  return(l0_c + l1_c*22 + l2_c*(22)^2 + l3_c*(22)^3 + alpha_c * 1875*price)
  } 
  # return((u_t-u_c)/abs(u_c)*100)
}

prices <- seq(from=0,to=1,by=.025)
ratios_c <- ratio_price(prices, treated = FALSE)
ratios_t <- ratio_price(prices, treated = TRUE)
data.frame(cbind(prices,Treatment=ratios_t, Control=ratios_c)) %>% 
  pivot_longer(
    cols      = c(Treatment, Control),
    names_to  = "Group",
    values_to = "ratio"
  ) %>%
  ggplot(aes(x=prices,y=ratio,color=Group,fill=Group)) + 
  geom_point() + 
  geom_vline(xintercept = 0.2, linetype='dashed') + 
  theme_minimal() + 
  labs(x="Cost Sharing %", y="Utilities", fill="Group")
```

## Counterfactuals

The process is: 
    1. Use equilibrium parameters for a given group ($T, C$) and a price scheme to estimate how many days a person stays in a SNF, conditional on entry. 
    2. This gives you an average expected valuation on intensive margin, $V(t)$
    3. This can be scaled by the price coefficient $\alpha$ to recover dollar values. 
    4. From there, unconditional averages are calculated by scaling on the extensive margin (0.01957 for the healthy group, 0.02457 for the sick group)
    5. Counterfactuals are run by varying the price schedule for each group.
    
```{r counterfactuals-old}
# Parameters to specify
load(paste0(mypath, "FullSimulations_20250415.RData", sep=""))
N <- 10000 # how many individuals to sample
treated <- TRUE # which group are we doing here?
myprice_pre <- 0 # vary this for counterfactuals (0, 37.5, 187.50)
myprice_post <- 37.5 # vary this for counterfactuals (0, 37.5, 187.50)
# sourceCpp(paste0(path, "simulate_person_value.cpp", sep=""), rebuild = TRUE) 
# sourceCpp(paste0(path, "simulate_person.cpp", sep=""), rebuild = TRUE) 

if (treated == TRUE) { 
  epscale_val <- epscale_t
  omega_val <- omega_t
  alpha_val <- alpha_t
  l0 <- l0_t
  l1 <- l1_t
  l2 <- l2_t
  l3 <- l3_t
} else { 
  epscale_val <- epscale_c
  omega_val <- omega_c
  alpha_val <- alpha_c
  l0 <- l0_c
  l1 <- l1_c
  l2 <- l2_c
  l3 <- l3_c
  }

# First, want to know how consumer max utility changes with price 
# make a dataset of N individuals with unique shocks (drawn once for all iterations)
simdata <- expand_grid(
  id = 1:N,
  t = 1:100
) %>%
  arrange(id, t) %>% 
  mutate(eps_k = epscale_val*revd(n = (100*N), scale=1), 
         eps_d = epscale_val*revd(n = (100*N), scale=1)) %>% 
    mutate(lambda = l0 + l1*t + l2*t^2 + l3*t^3) %>%
    mutate(p = ifelse(t < 22, myprice_pre, myprice_post),
           omega = omega_val,
           delta = 1,
           alpha = alpha_val,
           u = lambda + alpha*p )

# predict exit day for each person
preds <- data.frame(value=simulate_all_value_cpp(
  id     = simdata$id,
  u      = simdata$u,
  eps_d  = simdata$eps_d,
  eps_k  = simdata$eps_k,
  omega  = simdata$omega
), id=1:N, 
exit_day=simulate_all_cpp(
  id     = simdata$id,
  u      = simdata$u,
  eps_d  = simdata$eps_d,
  eps_k  = simdata$eps_k,
  omega  = simdata$omega
)) %>% filter(exit_day < 75) %>% # trim tails
  mutate(value = value/abs(alpha_val)) %>% # convert from utils to dollars 
  mutate(cost = exit_day * myprice_pre + pmax(exit_day-21, 0) * (myprice_post-myprice_pre)) %>% 
  mutate(netvalue = pmax(value - cost,0))

# unconditional value: 
if (treated == FALSE) { 
  preds <- preds %>% filter(netvalue > 0) %>% mutate(netvalue = netvalue * 0.01957)
  mean(preds$netvalue) # in base case, this is $24.74
    # with full insurance, this jumps up to $103.58 (318.7% increase)
   # at $37.50 every day this is $10.37
} else if (treated == TRUE) {
  preds <- preds %>% filter(netvalue > 0) %>% mutate(netvalue = netvalue * 0.02457)
  mean(preds$netvalue) # in base case, this is $190.18 (7.7 times higher)
    # with full insurance, this jumps up to $250.34 (31.6% increase)
  # at $37.50 every day this is $162.17
}

# Counterfactuals here are: 
```

### OLD CODE
```{r value-with-shocks, eval=FALSE, echo = FALSE}
timevars <- timevars %>% mutate(V_t = ifelse(t==100, u_t, NA),
                                V_c = ifelse(t==100, u_c, NA))

# transform shocks and calculate 
timevars <- timevars %>% mutate(eps_d_t = epscale_t*eps_d, 
                               eps_k_t = epscale_t*eps_k,
                               eps_d_c = epscale_c*eps_d, 
                               eps_k_c = epscale_c*eps_k)
for (d in 99:1) { # backward loop
  timevars <- timevars %>% 
    mutate(V_t = ifelse(t==d, u_t + pmax(eps_k_t + delta*lead(EV_t), eps_d_t), V_t),
           V_c = ifelse(t==d, u_c + pmax(eps_k_c + delta*lead(EV_c), eps_d_c), V_c))
}

# Plot
ggplot(timevars, aes(x=t)) + 
  geom_point(aes(y=V_t, color='Sick Spouse Group'), size=2) +
  geom_point(aes(y=V_c, color='Healthy Spouse Group'), size=2) +
  theme_minimal() + 
  labs(x = "Days in SNF", y = "", subtitle="Value With Shocks(t)", color=NULL) + 
  geom_hline(yintercept = 0) + 
  theme(legend.position="inside",
      legend.position.inside = c(0.01, 0.99),       # top-left corner inside the plot area
      legend.justification = c("left", "top"))
```

```{r exit-choice, eval=FALSE, echo=FALSE}
timevars <- timevars %>% mutate(cumV_t = cumsum(V_t),
                                cumV_c = cumsum(V_c))
ggplot(timevars, aes(x=t)) + 
  geom_point(aes(y=cumV_t, color='Sick Spouse Group'), size=2) +
  geom_point(aes(y=cumV_c, color='Healthy Spouse Group'), size=2) +
  theme_minimal() + 
  labs(x = "Days in SNF", y = "", subtitle="Cumulative Value(t)", color=NULL) + 
  geom_hline(yintercept = 0) + 
  theme(legend.position="inside",
      legend.position.inside = c(0.01, 0.99),       # top-left corner inside the plot area
      legend.justification = c("left", "top"))
```

